---
title: 'Assignment 3: Looping'
author: "Katheryn Moya"
date: "3/4/2023"
output: 
   html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    code_folding: show
    collapsed: yes
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(purrr)
library(ggpubr)
library(patchwork)
library(cowplot)
```

# Building a Highway 

For this analysis we will build a highway using four different car types, a range of speed data and the auto power function to calculate the power required to maintain various speeds 

* car A  mass 20000 kg, area 22 m^2^

* car B mass 65000 kg, area 30 m^2^

* car C mass 38000 kg area 22 m^2^

* car D mass 5000 kg and area 10m^2 *(a super light car)*

Mean highway speed of data is 80 km/hr with standard deviation of 10 km/hr

```{r sampling2, eval=TRUE, echo=TRUE}

# source function for auto power

source("~/Documents/MESM Year 2/ESM262/working/R/auto_power.R")

# set possible cars data frame
# add an additional super light car with mass 5,000 kg and a surface area 10m^2
possible_cars = data.frame(name = c("A","B","C", "D"),
                           mass = c(10000, 65000, 38000, 5000),
                           area = c(22, 30, 22, 10)) 

speed_base = 100 * 0.28 # designate a baseline for a mean speed of 100km/hr

# estimate power for each car type

possible_cars$power = auto_power(v=speed_base, A = possible_cars$area, m=possible_cars$mass)
```

## Sampling

```{r}
# create sample to get a distribution of speeds

nsample = 100

# assume a log normal distribution of speeds with mean 80km/hr, and a standard deviation of 10km/hr (12.5%)

mean_speed = log(80*0.277) # convert speed from m/s to km/hr # log normal distribution

speeds = rlnorm(mean = mean_speed, 
                sd = 0.125*mean_speed, 
                nsample) 
```

## Using *pmap* 

```{r}
power_pmap = pmap(list(A = possible_cars$area, 
                        m = possible_cars$mass),
                   auto_power, v = speeds)

# create dataframe with column names
power_pmap = as.data.frame(power_pmap, col.names= possible_cars$name)

longer_power_pmap<- power_pmap%>% 
  pivot_longer(cols=everything(), names_to="car", values_to="power") #pivot_longer to turn columns into rows
  
```

# For looping


```{r withforloop}
# create dataframe to store results 
power_loop = as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))

for (i in 1:4) {
  power_loop[,i] = auto_power(A = possible_cars$area[i], m = possible_cars$mass[i], v = speeds)
}

# add column names
colnames(power_loop)= possible_cars$name

longer_power_loop <- power_loop %>% 
  pivot_longer(cols = everything(), names_to="car", values_to ="power")

```



```{r}
# compare means between the two methods
power_pmap %>% map(mean)
power_loop%>% map(mean)

# upper 90% quantile
power_pmap %>% map(quantile, probs=0.9)
power_loop %>% map(quantile, probs=0.9)
```

## Data Visualization

```{r}
# quick visualization

pmap_power <- ggplot(data = longer_power_pmap,
                     aes(x = car, y = power)) +
              geom_boxplot(aes(fill = car)) +
              scale_fill_manual(values = c("cadetblue3", "darkseagreen3", "darkolivegreen3", "darkslategray4")) +
              theme_minimal() +
              labs(x = "Car Type",
                   y = "Power Required by Speed",
                   title = "Pmap Loop")

loop_power <- ggplot(data = longer_power_loop,
                     aes(x = car, y = power)) +
              geom_boxplot(aes(fill = car)) +
              scale_fill_manual(values = c("cadetblue3", "darkseagreen3", "darkolivegreen3", "darkslategray4")) +
              theme_minimal() +
              labs(x = "Car Type",
                   y = "Power Required by Speed",
                   title = "For Loop")
              

```


```{r, fig.width=9}
# put the plots together for comparison

plot_grid(pmap_power, loop_power, 
          align = 'h',
          greedy = TRUE,
          labels = c('A', 'B'), 
          label_size = 12,
          hjust = -0.5,
          vjust = 1.5) 

```